---
title: "Is bed a bottleneck"
author: "Linping"
date: "6/8/2018"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message=FALSE)
library(tidyverse)
library(lubridate)
library(zoo)
```

```{r include = FALSE}
adtLocation <- read_delim("/srv/rsdata/UCSD_ED_ADT_LOCATION.csv",
                          ";",
                          escape_double = FALSE,
                          col_types = cols(ADT_ARRIVAL_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                           IN_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                           OUT_DTTM = col_datetime(format = "%m/%d/%Y %H:%M")),
                          trim_ws = TRUE) %>% 
  filter(ADT_DEPARTMENT_NM_WID == 'EMERGENCY [100101010]') %>% 
  mutate(ADT_BED_LABEL_WID = gsub(' \\[.*\\]','', ADT_BED_LABEL_WID)) %>% 
  filter(!grepl('-WR', ADT_BED_LABEL_WID),
         !grepl('-EDX-', ADT_BED_LABEL_WID),
         !grepl('-TRG', ADT_BED_LABEL_WID),
         !grepl('-RMA', ADT_BED_LABEL_WID),
         !ADT_BED_LABEL_WID %in% c('zCHECKOUT', 'Flu Zone', 'CHARGER', 'EDS INTERNAL','SIM')) %>% 
  mutate(ADT_BED_LABEL_WID = plyr::mapvalues(ADT_BED_LABEL_WID, 
                                             from = c('12-FLU', '13-FLU','14-FLU','15-FLU','16-FLU','17-FLU','18-FLU','19-FLU','20-FLU','21-FLU','A- Y','01-Tr','02-Tr'),
                                             to = c('12-B','13-B','14-B','15-B','16-B','17-O','18-O','19-O','20-O','21-O','A-Y','01-Y','02-Y'))) %>% 
  select(PAT_ENC_CSN_ID, IN_DTTM, OUT_DTTM, ADT_BED_LABEL_WID, ADT_ARRIVAL_DTTM) %>% 
  gather(key =type, value = EVENT_TIME, -PAT_ENC_CSN_ID, -ADT_BED_LABEL_WID, -ADT_ARRIVAL_DTTM)

events <- read_delim("/srv/rsdata/UCSD_ED_EVENTS.csv",
                     ";",
                     escape_double = FALSE,
                     col_types = cols(ADT_ARRIVAL_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                      EVENT_TIME = col_datetime(format = "%m/%d/%Y %H:%M")),
                     trim_ws = TRUE)

encounters <- read_csv("/srv/rsdata/UCSD_F_ED_ENCOUNTERS.csv") %>% 
  mutate(ADT_ARRIVAL_DTTM = dmy_hms(ADT_ARRIVAL_DTTM),
         ADT_ARRIVAL_DATE = dmy_hms(ADT_ARRIVAL_DATE)) %>% 
  filter(ED_OR_UC_DEPARTMENT == 'EMERGENCY')

useful <- encounters %>% select(PAT_ENC_CSN_ID, ACUITY_LEVEL_C, BEHAVIORAL_DASHB, FIRST_CHIEF_COMPLAINT)

filtered <- events %>% 
  inner_join(useful, by = 'PAT_ENC_CSN_ID')
```


## Beds number calculation

```{r}
total <- filtered %>% left_join(adtLocation, by = c('PAT_ENC_CSN_ID', "ADT_ARRIVAL_DTTM", "EVENT_TIME"))
total_small <- total %>% select(PAT_ENC_CSN_ID, EVENT_TIME, type, ADT_BED_LABEL_WID) %>% drop_na() %>% unique()

total_small <- total_small %>% mutate(type = plyr::mapvalues(type, from = c("IN_DTTM", "OUT_DTTM"), to = c('in','out')), label = paste(ADT_BED_LABEL_WID, type, sep = '_'))
total_small <- total_small %>% mutate(date = date(EVENT_TIME), hour = hour(EVENT_TIME)) %>% arrange(date, hour)
total_new <- total_small %>% 
  dplyr::rename(bed = ADT_BED_LABEL_WID) %>% 
  select(date, hour, bed, type) %>% 
  dplyr::group_by(date, hour) %>% 
  dplyr::mutate(id = row_number()) %>% 
  spread(bed, type)
total_new <- total_new %>% replace(na.locf(total_new) == 'in', 'in')
total_new <- total_new %>% 
  select(-id) %>% 
  unique()
repl <- total_new[,3:65]
repl[repl == 'in'] <- 1
repl[repl == 'out'] <- 0
repl <- as.data.frame(sapply(repl, as.numeric))
total_new[,3:65] <- repl
```


```{r}
total_new$rowSum <- rowSums(total_new[3:65],na.rm=TRUE)

total_final <- total_new %>% select(date, hour, rowSum) %>% 
  group_by(date, hour) %>% 
  dplyr::summarise(numBeds = max(rowSum)) %>% 
  ungroup()
```

For example on 2014-01-01, from midnight to 1am , we labeled one transfer-in as 1, and transfer-out as 0 for each bed and sum across each bed. And we group by date and hour, and use the maximum number of that sum. we use this maximum number as the approximation for the number of beds used in that hour on that day

```{r}
DT::datatable(total_new %>% head(4), options=list(autoWidth=TRUE, scrollX=TRUE))
```

```{r}
hist(total_final$numBeds)
```

## Physician Number Calculation

To calculate the number of physicians appeared within each hour, we group by the name of physician and date. For a particular physician, we find out the time of both first and last show up time, and we assume that during the time period between fist and last show up time, the physician was working. We then calculate the census of all physician, and the result is the number of physicians working within each hour. 

The assumption of this calculation:

  Physician do not leave the hospital early in the day, and return late in the same day (i.e. go home at 1am and back to work at 11pm)

```{r}
filter_date <- seq(from = as.Date('2015-08-16'),
                   to = as.Date('2015-08-23'),
                   by = 'day')
f_ori <- filtered %>% 
  filter(PROV_TYPE == 'Physician') %>% 
  mutate(date = date(EVENT_TIME),
         hour = hour(EVENT_TIME)) %>% 
  mutate(in_date = date %in% filter_date,
         event = EVENT_NAME == 'Charting Complete',
         true = in_date&event) %>% 
  group_by(date,hour, PROV_NAME) %>% 
  mutate(n = n())
  
f <- f_ori %>% 
  filter(!true) %>%
  ungroup() %>% 
  group_by(date, PROV_NAME) %>% 
  summarise(first = min(EVENT_TIME),
            last = max(EVENT_TIME))

phy_census <- data.frame(
  time = seq(from = ymd_h('2014-01-02 0'),
             to = ymd_h('2017-12-31 23'),
             by = 'hour'))

for(i in 1:nrow(phy_census)){
  phy_census$count[i] <- sum(f$last > phy_census[i,'time'] & f$first <= phy_census[i,'time'], na.rm=TRUE)
}
phy_census_0 <- phy_census %>% mutate(date = date(time), hour = hour(time))


f <- f_ori %>% 
  filter(!true, n > 1) %>%
  ungroup() %>% 
  group_by(date, PROV_NAME) %>% 
  summarise(first = min(EVENT_TIME),
            last = max(EVENT_TIME))

for(i in 1:nrow(phy_census)){
  phy_census$count[i] <- sum(f$last > phy_census[i,'time'] & f$first <= phy_census[i,'time'], na.rm=TRUE)
}
phy_census_1 <- phy_census %>% mutate(date = date(time), hour = hour(time))


f <- f_ori %>% 
  filter(!true, n > 2) %>%
  ungroup() %>% 
  group_by(date, PROV_NAME) %>% 
  summarise(first = min(EVENT_TIME),
            last = max(EVENT_TIME))

for(i in 1:nrow(phy_census)){
  phy_census$count[i] <- sum(f$last > phy_census[i,'time'] & f$first <= phy_census[i,'time'], na.rm=TRUE)
}
phy_census_2 <- phy_census %>% mutate(date = date(time), hour = hour(time))

f <- f_ori %>%  
  filter(!true, n > 3) %>%
  ungroup() %>% 
  group_by(date, PROV_NAME) %>% 
  summarise(first = min(EVENT_TIME),
            last = max(EVENT_TIME))

for(i in 1:nrow(phy_census)){
  phy_census$count[i] <- sum(f$last > phy_census[i,'time'] & f$first <= phy_census[i,'time'], na.rm=TRUE)
}
phy_census_3 <- phy_census %>% mutate(date = date(time), hour = hour(time))

```

We believe this is currently the optimal approximation we can find based on the data we have. But we still find that, the number of physician seems to be higher than we expected. The maximum number is `r max(phy_census$count)`. And the distribution is as follows:

```{r}
hist(phy_census_0$count)
hist(phy_census_1$count)
hist(phy_census_2$count)
hist(phy_census_3$count)
```

** Possible improvement: ** Calculate time difference between previous removal from a patient and next assignment to another patient of one physician. But we need to figure out the threshold (length of time) of how to define a physician is not at work. 
  
## Patient Output Number

The number of patient output is defined by calculating how many patients encountered the event `Patient departed from ED`. 
```{r}
patient <- filtered %>% 
  filter(EVENT_NAME == 'Patient departed from ED') %>% 
  mutate(date = date(EVENT_TIME),
         hour = hour(EVENT_TIME)) %>% 
  group_by(date, hour) %>% 
  dplyr::summarise(patient = n())
```


## Summary

```{r}
summary <- total_final %>% 
  select(date, hour, numBeds) %>% 
  inner_join(phy_census_0, by = c('date','hour')) %>% 
  inner_join(patient, by = c('date','hour')) %>%
  select(-time) %>%
  rename(physician = count)

summary <- summary %>% 
  mutate(month = month(date),
         weekday = wday(date)) %>% 
  mutate(season = ifelse(month %in% c(1,2,3), "Winter", ifelse(month %in% c(6,7,8), 'Summer', 'Spring/Fall')))
```

```{r}
df <- data.frame(hour = rep(seq(0,23,1),3),
                 season = c(rep('Summer',24), rep("Spring/Fall", 24), rep('Winter', 24)),
                 beds = c(rep(19,7), rep(15,2),rep(23,5), rep(27,10),
                          rep(38,2), rep(26,2),rep(19,3),rep(27, 2),rep(30,3),rep(38,12),
                          rep(51,2), rep(34,2),rep(26,3),rep(34,2),rep(51,7), rep(59, 8)))
```

```{r}
hmm <- summary %>% left_join(df, by = c('season','hour')) %>% 
  mutate(utilization = numBeds / beds)
```

We also define utilization as $$\frac{\text{number of beds used}}{\text{total beds available}}$$. 

The bed schedule given to us cannot be perfectly matched with the calculation we've had. `r nrow(hmm %>% filter(utilization > 1))` out of `r nrow(hmm)` have number of beds observed greater than scheduled. 

The reasons may vary, we think:

  1. They may change their schedule based on their immediate judgement of the demand;
  2. Patients were still in bed for treatment so that the bed cannot be closed according to the schedule
  3. We need to improve the algorithm to estimate the number of beds used

Based on previous analysis, demand peak usually happens during 6pm through midnight. And the ED will open as many beds as possible during those hours too.

```{r}
output <- hmm %>% 
  filter(hour %in% c(18, 19, 20, 21, 22,23, 0), utilization > 0.5) %>% 
  group_by(physician) %>% 
  summarise(avg = mean(patient), numObservations = n())

output %>% 
  ggplot(aes(x = physician, y = avg, size = numObservations)) +
  ggtitle('Peak hour average output (all physicians count)') +
  geom_point(show.legend = FALSE)+
  geom_smooth(show.legend = FALSE) + 
  labs(x='number of physicians within given hour', y='Average patient output') +
  theme_bw()
```


```{r}
summary <- total_final %>% 
  select(date, hour, numBeds) %>% 
  inner_join(phy_census_1, by = c('date','hour')) %>% 
  inner_join(patient, by = c('date','hour')) %>%
  select(-time) %>%
  rename(physician = count)

summary <- summary %>% 
  mutate(month = month(date),
         weekday = wday(date)) %>% 
  mutate(season = ifelse(month %in% c(1,2,3), "Winter", ifelse(month %in% c(6,7,8), 'Summer', 'Spring/Fall')))

hmm <- summary %>% left_join(df, by = c('season','hour')) %>% 
  mutate(utilization = numBeds / beds)


output <- hmm %>% 
  filter(hour %in% c(18, 19, 20, 21, 22,23, 0), utilization > 0.5) %>% 
  group_by(physician) %>% 
  summarise(avg = mean(patient), numObservations = n())


output %>% 
  ggplot(aes(x = physician, y = avg, size = numObservations)) +
  ggtitle('Peak hour average output (physicians saw more than 1 patients count)') +
  geom_point(show.legend = FALSE)+
  geom_smooth(show.legend = FALSE) + 
  labs(x='number of physicians within given hour', y='Average patient output') +
  theme_bw()
```


```{r}
summary <- total_final %>% 
  select(date, hour, numBeds) %>% 
  inner_join(phy_census_2, by = c('date','hour')) %>% 
  inner_join(patient, by = c('date','hour')) %>%
  select(-time) %>%
  rename(physician = count)

summary <- summary %>% 
  mutate(month = month(date),
         weekday = wday(date)) %>% 
  mutate(season = ifelse(month %in% c(1,2,3), "Winter", ifelse(month %in% c(6,7,8), 'Summer', 'Spring/Fall')))

hmm <- summary %>% left_join(df, by = c('season','hour')) %>% 
  mutate(utilization = numBeds / beds)


output <- hmm %>% 
  filter(hour %in% c(18, 19, 20, 21, 22,23, 0), utilization > 0.5) %>% 
  group_by(physician) %>% 
  summarise(avg = mean(patient), numObservations = n())

output %>% 
  ggplot(aes(x = physician, y = avg, size = numObservations)) +
  ggtitle('Peak hour average output (physicians saw more than 2 patients count)') +
  geom_point(show.legend = FALSE)+
  geom_smooth(show.legend = FALSE) + 
  labs(x='number of physicians within given hour', y='Average patient output') +
  theme_bw()
```

```{r}
summary <- total_final %>% 
  select(date, hour, numBeds) %>% 
  inner_join(phy_census_3, by = c('date','hour')) %>% 
  inner_join(patient, by = c('date','hour')) %>%
  select(-time) %>%
  rename(physician = count)

summary <- summary %>% 
  mutate(month = month(date),
         weekday = wday(date)) %>% 
  mutate(season = ifelse(month %in% c(1,2,3), "Winter", ifelse(month %in% c(6,7,8), 'Summer', 'Spring/Fall')))

hmm <- summary %>% left_join(df, by = c('season','hour')) %>% 
  mutate(utilization = numBeds / beds)

output <- hmm %>% 
  filter(hour %in% c(18, 19, 20, 21, 22,23, 0), utilization > 0.5) %>% 
  group_by(physician) %>% 
  summarise(avg = mean(patient), numObservations = n())


output %>% 
  ggplot(aes(x = physician, y = avg, size = numObservations)) +
  ggtitle('Peak hour average output (physicians saw more than 3 patients count)') +
  geom_point(show.legend = FALSE)+
  geom_smooth(show.legend = FALSE) + 
  labs(x='number of physicians within given hour', y='Average patient output') +
  theme_bw()
```
