---
title: "Current Model Recreate"
output:
  html_document:
   theme: flatly
   highlight: tango
   code_folding: hide
---

Here is a link to get back to the original website [model.html](model.html)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      fig.width = 16,
                      fig.height = 9,
                      fig.align = 'center')
library(tidyverse)
library(lubridate)
```

**In this page, we will recreate part of the current model**

## Current Model Breakdown

**4 major components are considered in the current model: **

  * **Season:** *High, Medium, Low Demand Season* (Based on their understanding of average daily patient volume)
  
  * **Weekday:** *Sun/Mon, Tue-Sat* (Assuming Sunday and Monday have more patients than that of the other days of week)
  
  * **Acuities:** *Acuities1-2-3, Acuities4-5* (Industrial agreement of treating throughput: 2.15 pts/hour for acuities1-2-3, and 3.00 pts/hour for acuities4-5)
  
  * **Historical Volume** *75th percentile* of the patient volumes for a specific hour given season, weekday and acuitites (Estimated by tradeoffs between cost and benefits)


```{r include=FALSE}
Pts123 <- 2.15
Pts45 <- 3.00
serviceLevel <- 0.75
encounters <- read_csv("/srv/rsdata/UCSD_F_ED_ENCOUNTERS.csv",
                       col_types = cols(ADT_ARRIVAL_DATE = col_date(format = "%m/%d/%Y"),
                                        ADT_ARRIVAL_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                        ED_DEPARTURE_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                        ED_DISPOSITION_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                        EMERGENCY_ADMISSION_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                        HOSPITAL_ADMISSION_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                        HOSPITAL_DISCHARGE_DATE = col_date(format = "%m/%d/%Y"),
                                        HOSPITAL_DISCHARGE_DTTM = col_datetime(format = "%m/%d/%Y %H:%M"),
                                        UPDATE_DATE = col_datetime(format = "%m/%d/%Y %H:%M")))

encounters <- bind_rows(
  encounters %>% filter(ED_OR_UC_DEPARTMENT == 'EMERGENCY'),
  encounters %>% filter(ED_OR_UC_DEPARTMENT == 'Kiswahili') %>% 
    mutate(LANGUAGE = paste(LANGUAGE, ED_OR_UC_DEPARTMENT, sep = '/'),
           ED_OR_UC_DEPARTMENT = BEHAVIORAL_DASHB,
           BEHAVIORAL_DASHB = as.integer(X77)),
  encounters %>% filter(ED_OR_UC_DEPARTMENT == 'English') %>% 
    mutate(FINANCIAL_CLASS_NAME = PATIENT_SEX,
           PATIENT_SEX = LANGUAGE,
           LANGUAGE = ED_OR_UC_DEPARTMENT,
           ED_OR_UC_DEPARTMENT = BEHAVIORAL_DASHB,
           BEHAVIORAL_DASHB = as.integer(X77))
) %>% 
  select(-X77)
```


## Scheduling Part

Currently, the initial physicians' time schedule is made by Dr. Carstairs. The schedule is usually released 4 month prior, and is subjected to make slightly changes no latter than 2 weeks prior (but it's hard and require negotiation).

We, in this document, will use the plan for a high demand season, Monday/Sunday as a demonstration.
The schedule looks like below:


```{r include=FALSE}
highAcSchedule <- data_frame(Zone = c('Y1', 'Y2', 'Y1', 'Y2', 'Y1', 'Y2', 
                                      'Orange', 'Orange', 'Orange', 'Float'),
                             TimeSlot = c('6a-3p', '7a-4p', '2p-11p', 
                                          '3p-12a', '10p-6a', '11p-7a', 
                                          '6a-12p', '12p-7p', '7p-2a', '9a-3p'),
                             H0 = c(0,0,0,0,1,1,0,0,1,0),
                             H1 = c(0,0,0,0,1,1,0,0,1,0),
                             H2 = c(0,0,0,0,1,1,0,0,0,0),
                             H3 = c(0,0,0,0,1,1,0,0,0,0),
                             H4 = c(0,0,0,0,1,1,0,0,0,0),
                             H5 = c(0,0,0,0,1,1,0,0,0,0),
                             H6 = c(1,0,0,0,0,1,1,0,0,0),
                             H7 = c(1,1,0,0,0,0,1,0,0,0),
                             H8 = c(1,1,0,0,0,0,1,0,0,0),
                             H9 = c(1,1,0,0,0,0,1,0,0,1),
                             H10 = c(1,1,0,0,0,0,1,0,0,1),
                             H11 = c(1,1,0,0,0,0,1,0,0,1),
                             H12 = c(1,1,0,0,0,0,0,1,0,1),
                             H13 = c(1,1,0,0,0,0,0,1,0,1),
                             H14 = c(0,1,1,0,0,0,0,1,0,1),
                             H15 = c(0,0,1,1,0,0,0,1,0,1),
                             H16 = c(0,0,1,1,0,0,0,1,0,0),
                             H17 = c(0,0,1,1,0,0,0,1,0,0),
                             H18 = c(0,0,1,1,0,0,0,1,0,0),
                             H19 = c(0,0,1,1,0,0,0,0,1,0),
                             H20 = c(0,0,1,1,0,0,0,0,1,0),
                             H21 = c(0,0,1,1,0,0,0,0,1,0),
                             H22 = c(0,0,0,1,1,0,0,0,1,0),
                             H23 = c(0,0,0,0,1,1,0,0,1,0))


lowAcSchedule <- data_frame(Zone = c('FT am', 'FT pm', 'FT ON', 'P', 
                                     'P', 'EDS 1', 'EDS2'),
                            TimeSlow = c('9a-4p', '4p-10p', '9:30p-4a', 
                                         '10a-6p', '6p-2a', 
                                         '5:30p-1:30a(weekdays)','5p-12p'),
                            H0 = c(0,0,1,0,1,1,0),
                            H1 = c(0,0,1,0,1,.5,0),
                            H2 = c(0,0,1,0,0,0,0),
                            H3 = c(0,0,1,0,0,0,0),
                            H4 = rep(0,7),
                            H5 = rep(0,7),
                            H6 = rep(0,7),
                            H7 = rep(0,7),
                            H8 = rep(0,7),
                            H9 = c(1,0,0,0,0,0,0),
                            H10 = c(1,0,0,0,0,0,0),
                            H11 = c(1,0,0,0,0,0,0),
                            H12 = c(1,0,0,1,0,0,0),
                            H13 = c(1,0,0,1,0,0,0),
                            H14 = c(1,0,0,1,0,0,0),
                            H15 = c(1,0,0,1,0,0,0),
                            H16 = c(0,1,0,1,0,0,0),
                            H17 = c(0,1,0,1,0,.5,1),
                            H18 = c(0,1,0,0,1,1,1),
                            H19 = c(0,1,0,0,1,1,1),
                            H20 = c(0,1,0,0,1,1,1),
                            H21 = c(0,1,.5,0,1,1,1),
                            H22 = c(0,0,1,0,1,1,1),
                            H23 = c(0,0,1,0,1,1,1))

```

### Sample Table

```{r}
DT::datatable(highAcSchedule, caption = "Sample Schedule for Acuities 1-2-3", options = list(autoWidth=TRUE, scrollX=TRUE))
DT::datatable(lowAcSchedule, caption = "Sample Schedule for Acuities 4-5", options = list(autoWidth=TRUE, scrollX=TRUE))
```

## 75th Percentile
```{r include=FALSE}
data <- encounters %>% 
  # filter(ACUITY_LEVEL_C != 'null') %>% # when this filter applies, the number seems to match the spreadsheet model;
  mutate(ADT_WEEKDAYS = factor(wday(ADT_ARRIVAL_DTTM),
                               labels = c('Monday','Tuesday','Wednesday','Thursday','Friday','Saturday','Sunday')),
         ADT_MONTH = factor(month(ADT_ARRIVAL_DTTM),
                            labels = c('January','February','March','April','May','June','July','August','September','October','November','December')),
         ACUITY = factor(ifelse(ACUITY_LEVEL_C %in% c('1','2','3'), 'High', 'Low'),
                         levels = c('Low', 'High')),
         ADT_YEAR = factor(year(ADT_ARRIVAL_DTTM)),
         ADT_HOUR = factor(hour(ADT_ARRIVAL_DTTM)),
         SEASON = factor(ifelse(ADT_MONTH %in% c('January', 'February', 'March'), 'High', 
                                ifelse(ADT_MONTH %in% c('June', 'July', 'August'), 'Low', 'Medium'))),
         WEEKDAY = factor(ifelse(ADT_WEEKDAYS %in% c('Sunday', 'Monday'), 'Mon/Sun', 'Tue/Sat')))
```

### Brief exploration

```{r}
PtCount <- data %>% 
  group_by(ADT_ARRIVAL_DATE, ADT_HOUR, ACUITY, SEASON, WEEKDAY) %>% 
  summarise(numPt = n()) %>% 
  ungroup() %>% 
  filter(!is.na(ADT_HOUR)) %>% 
  ggplot(aes(x = ADT_HOUR, y = numPt, color = ACUITY)) +
  geom_boxplot() +
  facet_grid(SEASON~WEEKDAY) +
  labs(x = NULL, y = NULL) +
  ggtitle('Distribution of Patients count') +
  theme_bw() +
  scale_color_brewer(palette = 'Dark2') +
  theme(axis.text.x = element_text(angle = 90),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(linetype = 2),
        legend.position = 'bottom',
        legend.title = element_text(size = 8),
        axis.ticks = element_line())
PtCount
```

### Calculation

We use the following code to calculate the 75th percentile of patient volume by season, weekday, acuities and hour.

  * `ADT_ARRIVAL_DATE`: patient arrival date
  * `ADT_HOUR`: patient arrival hour
  * `ACUITY`: Acuities1-2-3 are coded as 'High'; Acuities4-5 are coded as 'Low'
  * `SEASON`: January and Feburary are coded as 'High'; June, July and August are coded as 'Low'; others are coded as 'Medium'
  * `WEEKDAY`: Grouped by Mon/Sun and Tue-Sat


```{r echo=TRUE}
PtVolume <- data %>% 
  group_by(ADT_ARRIVAL_DATE, ADT_HOUR, ACUITY, SEASON, WEEKDAY) %>% 
  summarise(numPt = n()) %>% 
  ungroup() %>% 
  group_by(ADT_HOUR, ACUITY, SEASON, WEEKDAY) %>% 
  summarise(numPercentile = quantile(numPt, serviceLevel)) %>% 
  ungroup()
DT::datatable(PtVolume, options = list(autoWidth=TRUE))
```

### Recreate 75th Percentile Plot

#### **Mon/Sun**

```{r}
MonSun <- PtVolume %>% 
  filter(SEASON == 'High',
         # ACUITY == 'High',
         WEEKDAY == 'Mon/Sun') %>% 
  ggplot(aes(x = ADT_HOUR, y = numPercentile, fill = ACUITY)) +
  geom_col(position = 'dodge') +
  theme_bw() + 
  ggtitle('ED Arrivals 75th Percentile - Sun-Mon - High Demand Season') +
  labs(x = NULL, y = NULL) +
  theme(legend.position = 'bottom',
        panel.grid.major.x = element_blank(),
        legend.title = element_text(size = 8)) +
  scale_fill_brewer(palette = 'Set2') +
  scale_y_continuous(limits = c(0,15), breaks = seq(0,15,3))
MonSun
```

#### **Tue/Sat**

```{r}
TueSat <- PtVolume %>% 
  filter(SEASON == 'High',
         # ACUITY == 'High',
         WEEKDAY == 'Tue/Sat') %>% 
  ggplot(aes(x = ADT_HOUR, y = numPercentile, fill = ACUITY)) +
  geom_col(position = 'dodge') +
  theme_bw() + 
  ggtitle('ED Arrivals 75th Percentile - Tue-Sat - High Demand Season') +
  labs(x = NULL, y = NULL) +
  theme(legend.position = 'bottom',
        panel.grid.major.x = element_blank(),
        legend.title = element_text(size = 8)) +
  scale_fill_brewer(palette = 'Set2') +
  scale_y_continuous(limits = c(0,15), breaks = seq(0,15,3))
TueSat
```

## Model Recreation
```{r include=FALSE}
# staff care need calculation
calcStaff <- function(df){
  smooth <- df$smooth
  smooth <- c(smooth[23:24], smooth)
  staff <- rep(NA, 26)
  
  for(i in 1:length(smooth)){
    staff[i] <-max(smooth[i], smooth[i+1], smooth[i+2])
  }
  return(staff[1:24])
}
```

### Calculation

`smooth` represents `Patient Arrivals Estimate`, is calculated as follows:

  * 24 hours are smoothed by 8 3-hour groups, (0-2, 3-5, etc.).
  * Average three 75th percentile patient volumes and use the average number for all hours within each 3-hour group.

`staffingNeed` is calculated as follows:
  
  * current hour cell denote as $h_0$, 
  * select the maximum `Patient Arrivals Estimate` among $h_0$, $h_{-1}$, and $h_{-2}$
  * divide the maximum by the common throughput by acuities

### Code
*The following code shows the calculation step*

```{r echo=TRUE}
datPassH <- PtVolume %>% 
  filter(SEASON == 'High',
         ACUITY == 'High',
         WEEKDAY == 'Mon/Sun') %>% 
  select(ADT_HOUR, numPercentile) %>% 
  mutate(group = rep(1:8, each = 3)) %>% 
  group_by(group) %>% 
  mutate(smooth = mean(numPercentile)) %>% 
  ungroup() %>% 
  select(ADT_HOUR, smooth) 
datPassH <- datPassH %>% 
  mutate(staffingNeed = calcStaff(datPassH) / Pts123)

datPassL <- PtVolume %>% 
  filter(SEASON == 'High',
         ACUITY == 'Low',
         WEEKDAY == 'Mon/Sun') %>% 
  select(ADT_HOUR, numPercentile) %>% 
  mutate(group = rep(1:8, each = 3)) %>% 
  group_by(group) %>% 
  mutate(smooth = mean(numPercentile)) %>% 
  ungroup() %>% 
  select(ADT_HOUR, smooth)
datPassL <- datPassL %>% 
  mutate(staffingNeed = calcStaff(datPassL) / Pts45)
```

### Recreation

We then combine the two together to get the final presentable model

```{r}
df <- data_frame(hour = factor(seq(0,23,1)), 
                 patientEstimateH = datPassH$smooth,
                 staffingNeedH = datPassH$staffingNeed,
                 staffingPlanH = colSums(highAcSchedule[-1:-2], na.rm = TRUE),
                 patientEstimateL = datPassL$smooth,
                 staffingNeedL = datPassL$staffingNeed,
                 staffingPlanL = colSums(lowAcSchedule[-1:-2], na.rm = TRUE),
                 staffingDiffH = - staffingNeedH + staffingPlanH,
                 staffingDiffL = - staffingNeedL + staffingPlanL,
                 totalPtEst = patientEstimateH + patientEstimateL,
                 totalStaffingNeed = staffingNeedH + staffingNeedL,
                 totalStaffingPlan = staffingPlanH + staffingPlanL,
                 totalStaffingDiff = - totalStaffingNeed + totalStaffingPlan)

DT::datatable(df, options = list(autoWidth=TRUE, scrollX=TRUE)) %>% DT::formatRound(columns = colnames(df)[-1])
```

### Useful Plots

**Recreate Staffing planned vs. Staffing Est Plot**

#### High Acuities

```{r}
highAcuities <- df %>% ggplot(aes(x = hour)) +
  geom_col(aes(y = staffingPlanH), fill = 'blue', alpha = 0.5) +
  geom_errorbar(aes(ymax = staffingNeedH, ymin = staffingNeedH)) +
  geom_text(aes(y = staffingNeedH, label = round(staffingNeedH, 2)), vjust = -0.5, size = 5) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  scale_fill_brewer(palette = 'Set2') +
  scale_color_brewer(palette = 'Set2') +
  scale_y_continuous(limits = c(0,12),
                     breaks = seq(0,12,2))
highAcuities
```


#### Low Acuities

```{r}
lowAcuities <- df %>% ggplot(aes(x = hour)) +
  geom_col(aes(y = staffingPlanL), fill = 'blue', alpha = 0.5) +
  geom_errorbar(aes(ymax = staffingNeedL, ymin = staffingNeedL)) +
  geom_text(aes(y = staffingNeedL, label = round(staffingNeedL, 2)), vjust = -0.5, size = 5) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  scale_fill_brewer(palette = 'Set2') +
  scale_color_brewer(palette = 'Set2') +
  scale_y_continuous(limits = c(0,12),
                     breaks = seq(0,12,2))
lowAcuities
```

#### Total

```{r}
total <- df %>% ggplot(aes(x = hour)) +
  geom_col(aes(y = totalStaffingPlan), fill = 'blue', alpha = 0.5) +
  geom_errorbar(aes(ymax = totalStaffingNeed, ymin = totalStaffingNeed)) +
  geom_text(aes(y = totalStaffingNeed, label = round(totalStaffingNeed, 2)), vjust = -0.5, size = 5) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  scale_fill_brewer(palette = 'Set2') +
  scale_color_brewer(palette = 'Set2') +
  scale_y_continuous(limits = c(0,12),
                     breaks = seq(0,12,2))
total
```

