---
title: "Sand Diego County Flu"
output: 
  html_document:
   theme: flatly
   highlight: tango
   code_folding: hide
---

<style>
pre {
  width: 560px;
  white-space: pre;
  word-wrap: normal;
  background-color: #ffffff;
}
</style>

```{r setup, include = FALSE}
library(tidyverse)
library(radiant)
library(DT)
# options(scipen = 100, width = 250)
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE,
  message = FALSE, comment = NA,
  fig.width = 16, fig.height = 9, fig.align = 'center'
)
```

```{r include=FALSE}

sdcounty <- readxl::read_xlsx("data/Flu Cases by Episode Week County of San Diego 04182018.xlsx",col_names = TRUE)
names(sdcounty) <- sdcounty[1,]
colnames(sdcounty)[colnames(sdcounty) == 'Season'] <- "Week"
sdcounty <- sdcounty[3:55,] 
gathered <- gather(sdcounty,Season,value=influenza,-Week)
gathered$Week <- as.integer(gathered$Week)
gathered$influenza <- as.numeric(gathered$influenza)
gathered$Year <- c(rep(2017,27),rep(2018,26),rep(2016,27),rep(2017,26),rep(2015,27),rep(2016,26),rep(2014,27),rep(2015,26),rep(2013,27),rep(2014,26),rep(2012,27),rep(2013,26),rep(2011,27),rep(2012,26),rep(2010,27),rep(2011,26),rep(2009,27),rep(2010,26),rep(2008,27),rep(2009,26),rep(2007,27),rep(2008,26),rep(2006,27),rep(2007,26),rep(2005,27),rep(2006,26),rep(2004,27),rep(2005,26),rep(2003,27),rep(2004,26))
recentsixyears <- gathered %>% 
  filter(Year %in%  c("2018","2017","2016","2015","2014","2013","2012"))
```

## {.tabset}

### General Introduction

The San Diego Flu Cases dataset is prepared by Epidemiology and Immunization Services Branch on Apr.18, 2018. It records the number of influenza cases reported every week from 2003-2017. This data will help us predict weekly epidemic situation and thus predict the weekly demand of Rady Children's Hospital.

For more detailed influenza cases report, please go to https://www.sandiegocounty.gov/content/sdc/hhsa/programs/phs/community_epidemiology/disease-surveillance.html.

The County's influenza surveillance program is a collaborative effort between San Diego County Public Health Services' Epidemiology & Immunization Services Branch, our medical community and laboratory partners, and the [San Diego Health Connect](http://sdhealthconnect.org/) Health Information Exchange.

Contact us at EpiDiv.HHSA@sdcounty.ca.gov or (619)692-8499 with any questions. Thank you for your interest.

```{r}
DT::datatable(sdcounty, options = list(scrollX = TRUE, autoWidth = TRUE))
```

* Cases are grouped by CDC Week using Episode Date, which is the date of onset, or the earliest date in which the case is identified.

### Recent Six Year's Trend

```{r warning=FALSE}
ggplot(recentsixyears,mapping = aes(x=as.numeric(Week),y=influenza,color=as.factor(Year)))+
  geom_point(show.legend = FALSE)+
  geom_line(show.legend = FALSE)+
  facet_grid(Year~.)+
  ggtitle("Trend of Number of Inflenza Cases in Recent Six Years") +
  labs(x = 'Week', y = 'Influenza') +
  theme_bw() +
  theme(panel.grid.major = element_blank()) +
  scale_x_continuous(breaks = seq(1,53,4))
  # scale_color_brewer(palette = 'Set2')
```
The time period that flu comes has changed during the past six years. That change might be used in the time-series model.

### Flu Outbreak in Season 2017/2018

```{r include=FALSE}
result <- pivotr(
  gathered, 
  cvars = "Year", 
  nvar = "influenza", 
  fun = "max_rm", 
  nr = 15
)
summary(result)
# dtab(result, pageLength = -1) %>% render()
maxvalue <- result$tab; register("maxvalue")
```

```{r}
visualize(
  maxvalue, 
  xvar = "Year", 
  yvar = "influenza", 
  type = "bar", 
  fill = "Year", 
  custom = TRUE
)+
  geom_text(aes(x = Year, y = influenza, label = round(influenza, 2)), vjust = -1) +
  labs(title = 'Mean Number of Influenza Cases, County of San Diego',y = "Max Number of Inluenza Cases", x = "Year") + 
  theme_bw() +
  # scale_fill_brewer(palette = 'Set2') +
  theme(panel.grid.major = element_blank(),
        legend.position = "none")
```


```{r include=FALSE}
result <- pivotr(
  gathered, 
  cvars = "Year", 
  nvar = "influenza", 
  data_filter = "Year != 2018", 
  nr = 15
)
summary(result)
# dtab(result, pageLength = -1) %>% render()
meanvalue <- result$tab; register("meanvalue")
```

```{r}
visualize(
  meanvalue, 
  xvar = "Year", 
  yvar = "influenza", 
  type = "bar", 
  fill = "Year", 
  custom = TRUE
)+
  geom_text(aes(x = Year, y = influenza, label = round(influenza, 2)), vjust = -1) +
  labs(title = 'Mean Number of Influenza Cases, County of San Diego',y = "Mean Number of Inluenza Cases", x = "Year") + 
  theme_bw() +
  # scale_fill_brewer(palette = 'Set2') +
  theme(panel.grid.major = element_blank(),
        legend.position = "none")
```


