---
title: Vacation Time 
output: 
  html_document:
   theme: flatly
   highlight: tango
   code_folding: hide
---

<style>
pre {
  width: 560px;
  white-space: pre;
  word-wrap: normal;
  background-color: #ffffff;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
library(tidyverse)
library(lubridate)
library(stringr)
library(knitr)
library(kableExtra)
```

Since patients go to Rady Children's Hospital Emergency Room will be students, students' activities should be relevent to the chance of going to Emergency Room. Thus, we collected data of when students take vacations.  

In the assumption, all students among San Diego will have the same schedule for vacations.

```{r}
vacation <- data.frame(Event = c("SpringBreakS", "SpringBreakE", "SummerStart","SummerEnd", "WinterStart", "WinterEnd",
                                 "v1", "v2", "v3", "v4", "v5", "v6", "v7", "v8", "v9", "v10", "thankss","thankse"),
                       Time = c("04-09", "04-20","06-13", "08-28","12-25", "01-05","01-15","02-16","02-19","02-20","02-21","05-28","07-04","09-04","09-22","11-10","11-20","11-24"))
```

```{r}
DT::datatable(vacation, options = list(autoWidth = TRUE))
# kable(vacation, "html", align = "c") %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed",  "bordered")) %>%
#   kable_styling() %>%
#   scroll_box(width = "100%", height = "500px")
```

The database contains information on:

* Event: the event when student take the break
* Time: month-date, time student are in the break. 





