---
title: "San Diego Weather"
output: 
  html_document:
   theme: flatly
   highlight: tango
   code_folding: hide
---

<style>
pre {
  width: 560px;
  white-space: pre;
  word-wrap: normal;
  background-color: #ffffff;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, 
  comment = NA, dpi = 144
)

library(radiant)
library(dplyr)
library(readr)
library(lubridate)

col <- c('year','month','day','hour','minute','temperature','relativeHumidity','meanSeaLevelPressure','totalPrecipitation','totalCloudCover','sunshineDuration','windSpeed','windDirection')
weather <- read_csv("data/history_export.csv", col_names = col, skip = 1)

options(repos="https://cran.rstudio.com" )
```

## Description

Data source: [www.meteoblue.com](https://www.meteoblue.com/en/weather/archive/export/san-diego_united-states-of-america_5391811)

`Relative Humidity`:  The amount of water vapor present in air expressed as a percentage of the amount needed for saturation at the same temperature.

`Mean Sea Level Pressure`:  The average atmospheric pressure at sea level.

`Total Precipitation`:  Total product of the condensation of atmospheric water vapor that falls under gravity. 

`Total Cloud Cover`:  The total fraction of the sky obscured by clouds when observed from a particular location.

`Sunshine Duration`:  Duration of sunshine in given period (usually, a day or a year) for a given location on Earth. 

`Wind Speed`:  Wind flow velocity (a fundamental atmospheric quantity).

`Wind Direction`:  The direction from which the wind originates.


## Weather Data
```{r}
DT::datatable(weather[1:500, c(-5)], options = list(autoWidth = TRUE, scrollX = TRUE))
```

<br>

## Weather Trend {.tabset}

### Humidity

```{r fig.width = 16, fig.height = 9, dpi = 400}
weather %>% 
  mutate(date = ymd(paste(year, month, day, sep = '-'))) %>% 
  group_by(date) %>% 
  summarise(avg = mean(relativeHumidity)) %>% 
  ggplot(aes(x = date, y = avg)) +
  geom_line(alpha = 0.8, size = 0.5) +
  labs(x = NULL, y = 'Daily Humidity Average') +
  theme_bw() +
  scale_x_date(date_breaks = '1 month') +
  theme(panel.grid.major = element_blank(),
        axis.text.x = element_text(angle = 90))

weather %>% ggplot(aes(x = factor(month), y = relativeHumidity)) +
  geom_boxplot() +
  theme_bw() +
  theme(panel.grid.major = element_blank()) +
  labs(x = 'Month', y = 'Relative Humidity')
```

### Temperature

```{r fig.width = 16, fig.height = 9, dpi = 400}
weather %>% 
  mutate(date = ymd(paste(year, month, day, sep = '-'))) %>% 
  group_by(date) %>% 
  summarise(avg = mean(temperature)) %>% 
  ggplot(aes(x = date, y = avg)) +
  geom_line(alpha = 0.8, size = 0.5) +
  # geom_hline(yintercept = mean(weather$temperature), size = 1, linetype = 2, color = 'red', alpha = 0.7) +
  labs(x = NULL, y = 'Daily Temperature Average') +
  theme_bw() +
  scale_x_date(date_breaks = '1 month') +
  theme(panel.grid.major = element_blank(),
        axis.text.x = element_text(angle = 90))

weather %>% ggplot(aes(x = factor(month), y = temperature)) +
  geom_boxplot() +
  theme_bw() +
  theme(panel.grid.major = element_blank()) +
  labs(x = 'Month', y = 'Temperature')
```
